\textls[-0]{\thetitle}

### 2023-03-16 canonicitymodel.tex
Save Canonicity Model.

Like most logical relation proof,  we interpret each judgement and inductively interpret each syntax piece. We are working on an intrinsic setting thus even if we omit context for brevity, the syntax piece is still well-typed. However, unlike most logical relation proof, we are having a proof-relevant logical realtion, essential for modeling universes properly~\cite{coquand2018canonicity}. Our canonicity model at MLTT part is basically the same as shown in \citet{coquand2018canonicity}.
% We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model). Interested audience can refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} for a complete technical treatements.

@@

\newcommand{\Glued}[1]{\ensuremath{{#1}^\bullet}}
\newcommand{\GluedPi}[2]{\ensuremath{\Pi^\bullet({#1},{#2})}}
\newcommand{\GSubstExt}[2]{\ensuremath{{#1},^\bullet{#2}}}
\newcommand{\Gpair}[2]{\ensuremath{({#1},^\bullet{#2})}}
\newcommand{\Gfst}[1]{\ensuremath{\texttt{fst}^\bullet~{#1}}}
\newcommand{\Gsnd}[1]{\ensuremath{\texttt{snd}^\bullet~{#1}}}
\newcommand{\Gsub}[2]{\ensuremath{{#1}\!\left[{#2}\right]^\bullet}}
\newcommand{\Glam}[1]{\ensuremath{\lambda^\bullet({#1})}}
\newcommand{\Gapp}[1]{\ensuremath{\texttt{app}^\bullet({#1})}}
\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^\bullet)^{#1}}}
\newcommand{\GLSigAdd}[3]{\ensuremath{\nu^{+\bullet}({#1},{#2},{#3})}}
\newcommand{\GCaseSig}[3]{\ensuremath{\texttt{CaseTy}^\bullet({#1},{#2},{#3})}}
\newcommand{\Gmodel}[1]{\ensuremath{{({#1})}^c}}

@@

Basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$; and for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$. Bascially the four sets are collecting the glued interpretation.

@@

\newcommand{\CWmodel}{\ensuremath{\mathit{W}^C}}
\newcommand{\CWsup}{\ensuremath{\mathit{Wsup}^C}}
\newcommand{\CWrec}{\ensuremath{\mathit{Wrec}^C}}

@@

  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{R}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{R}} \\

@@

  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}} &= \\

@@

\begin{align*}
  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\
    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \sub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) }{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}) \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
\end{align*}
... and its eliminator principle $\CWrec$. 
\begin{align*}
  \CWrec &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
    (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
    & \to [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\ 
    & \quad \quad \to 
    b \in (\GTm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{\SubstId}{{a}}}) }{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}) \\ 
    & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \\
    & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
  \CWrec & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda )
\end{align*}

Note that in $\CWmodel$, the $\Glued{b}$ uses the definition of $\denotesC{\wcode{\wsig}}$, which after unfolding, will refers to $\CWmodel$ again and in strictly positive position. So it is a well-founded definition.

>>>>

Like most logical relation proof,  we interpret each judgement and inductively interpret each syntax piece. We are working on an intrinsic setting thus even if we omit context for brevity, the syntax piece is still well-typed. However, unlike most logical relation proof, we are having a proof-relevant logical realtion, essential for modeling universes properly~\cite{coquand2018canonicity}. Our canonicity model at MLTT part is very close to the one shown in \citet{coquand2018canonicity} and \citet{sterling2019algebraic}.
% We will omit the treatment of the universe and the verification of most equality for simplicity (and the treatment of the universe will not affect our model). Interested audience can refer to \citet{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} for a complete technical treatements.

@@

\newcommand{\Glued}[1]{\ensuremath{{#1}^\bullet}}
\newcommand{\GluedPi}[2]{\ensuremath{\Pi^\bullet({#1},{#2})}}
\newcommand{\GSubstExt}[2]{\ensuremath{{#1},^\bullet{#2}}}
\newcommand{\Gpair}[2]{\ensuremath{({#1},^\bullet{#2})}}
\newcommand{\Gfst}[1]{\ensuremath{\texttt{fst}^\bullet~{#1}}}
\newcommand{\Gsnd}[1]{\ensuremath{\texttt{snd}^\bullet~{#1}}}
\newcommand{\Gsub}[2]{\ensuremath{{#1}\!\left[{#2}\right]^\bullet}}
\newcommand{\Glam}[1]{\ensuremath{\lambda^\bullet({#1})}}
\newcommand{\Gapp}[1]{\ensuremath{\texttt{app}^\bullet({#1})}}
\newcommand{\GSubstWeak}[1]{\ensuremath{(\texttt{p}^\bullet)^{#1}}}
\newcommand{\GLSigAdd}[3]{\ensuremath{\nu^{+\bullet}({#1},{#2},{#3})}}
\newcommand{\GCaseSig}[3]{\ensuremath{\texttt{CaseTy}^\bullet({#1},{#2},{#3})}}
\newcommand{\Gmodel}[1]{\ensuremath{{({#1})}^c}}
\newcommand{\GEl}[1]{\ensuremath{\Glued{\texttt{El}}({#1})}}
\newcommand{\Gwcode}[1]{\ensuremath{\Glued{\texttt{W}}({#1})}}
\newcommand{\GSubstId}{\ensuremath{\GSubstWeak{0}}}

@@

These four sets are collecting the glued interpretation:
basically the point is, for each well-typed type $\goodType{\Gm}{T}{j}$,  we have its denotation $\denotesCC{T} \in \GTy{j}{\denotesCC{\Gm}}$; and for each well-typed term $\goodTerm{\Gm}{t}{T}$, we have its denotation $\denotesCC{t} \in \GTm{\denotesCC{\Gm}}{\denotesCC{T}}$.

@@

\newcommand{\CWmodel}{\ensuremath{\mathit{W}^C}}
\newcommand{\CWsup}{\ensuremath{\mathit{W^Csup}}}
\newcommand{\CWrec}{\ensuremath{\mathit{W^Crec}}}

@@

  \denotesC{\goodType{\Gm}{\RecSig{\wsig}{"R"}}{}} &= \Gmodel{RS~\denotesC{\tau}~\denotesC{"R"}} \\

@@

  \denotesC{\goodTerm{\Gm}{\Wrec{\wsig}{\lkg}{t}}{T}}(\gm)(\gm') &= \CWrec~(\lambda w. \denotesC{"R"}(\gm)(\gm')(\Wrec{\wsig}{\sub{\lkg}{\gm'}}{w}))~f^r~\sub{t}{\gm'}~(\denotesC{t}(\gm)(\gm')) \\
  \text{ where } \quad f^r~j~\Glued{a}~b~b^C &= 
  \texttt{let}~\Glued{\rho}  \in \GTm{(\Glued{\cdot} \Glued{,} \Gsub{\Glued{B}}{\GSubstExt{\Glued{"id"}}{\Glued{a}}})}{(\Gsub{\Gsub{\Glued{R}}{\Glued{\gm}}}{\GSubstWeak{1}})} 
  \\
  & \ \texttt{ s.t.}~\Glued{\rho} \coloneq (\Wrec{\wsig}{\sub{\lkg}{\SubstComp{\gm}{\SubstWeak{1}}}}{b},b^C)~\texttt{ in } \\
  & \Gmodel{\Gsub{\Gapp{\Gsub{\Gapp{\Recproj{j}{\sub{\denotesCC{\lkg}}{\Glued{\gm}}}}}{\GSubstExt{\GSubstId}{\Glued{a}}}}}{\GSubstExt{\GSubstId}{\Glam{\Glued{\rho}}}}}(\epsilon)(\star) \\
  & \quad 
   \text{ and also } \goodTerm{\Gm}{\lkg}{\RecSig{\wsig}{"R"}} \quad ,\Glued{R} = \denotesCC{"R"}, \Glued{B} = \wsigproj{j}{2}{\Glued{\wsig}}

@@

\begin{align*}
  "Inductive"~\CWmodel &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{i + 1} \quad "where" \\
    \CWsup &: j < N \to \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\
    & \to \Glued{b} \in (\GTm{(\Glued{\cdot} \Glued{,} \Gsub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\GSubstId}{\Glued{a}}}) }{\Gsub{\GEl{\Gwcode{\wsig}}}{\GSubstWeak{1}}}) \to \CWmodel~\Glued{\wsig}~\Wsup{j}{\wsig}{a}{b}
\end{align*}
... and its eliminator principle $\CWrec$. 
\begin{align*}
  \CWrec \  &: (\Glued{\wsig} \in \GWSig{i}{\denotesCC{\cdot}}{N}) \to 
    (P : \{ t : \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \} \to \Set{k})  \\
    & \to [j < N \to  \Glued{a} \in \GTm{\Glued{\cdot}}{\wsigproj{j}{1}{\Glued{\wsig}}} \\ 
    & \quad \quad \to 
    b \in (\GTm{({\cdot} {,} \sub{\wsigproj{j}{2}{{\wsig}}}{\SubstExt{\SubstId}{{a}}}) }{\sub{\El{\wcode{\wsig}}}{\SubstWeak{1}}}) \\ 
    & \quad \quad \to b^C : (\Glued{\gm} \in (\Glued{\cdot} \Glued{,} \sub{\wsigproj{j}{2}{\Glued{\wsig}}}{\SubstExt{\SubstId}{\Glued{a}}}) \to P~\sub{b}{\Gmodel{\Glued{\gm}}}) \\
    & \quad \quad \to P~(\Wsup{j}{\wsig}{a}{b}) ] \\
    & \to \goodTerm{\cdot}{t}{\El{\wcode{\wsig}}} \to \CWmodel~\Glued{\wsig}~t \to P~t \\
  \CWrec \  & \Glued{\wsig}~P~f~t~(\CWsup~\Glued{a}~\Glued{b}) = f~\Glued{a}~b~(\lambda \Glued{\gm}. \CWrec~P~f~(\sub{b}{\Gmodel{\Glued{\gm}}})~(b^C~\Glued{\gm}))
\end{align*}

Note that in $\CWmodel$, the $\Glued{b}$ uses the definition of $\denotesC{\wcode{\wsig}}$, which after unfolding, will refers to $\CWmodel$ again and in strictly positive position. So it is a well-founded definition. We also don't distinguish $(b, b^C)$ and $\Glued{b}$ for simplicity.
